---
title: "Prediction"
author: "Xiaoyang Li"
date: "2020/11/5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(caret)
library(rpart)
library(rpart.plot)
library(pROC)
library(ggstatsplot)
```

## Data clean
```{r}

# data import
data = read_delim("data.dat", "\t", escape_double = FALSE, trim_ws = TRUE) 

# 96 patients and remove comments column
data = data[1:96,] %>% dplyr::select(SurgPathFindings,SeqentialStaging, Largestdiameterofappendixmeasuredanteriortoposteriorwithcompre_A, Maximalmuralthicknessdistancefromhyperechoicluminalinterfaceto_A,  Presenceoffecalith_A, Presenceofsurroundingfreefluid_A) %>% 
  rename(Surg = SurgPathFindings,
         US2 = SeqentialStaging,
         Appdiam = Largestdiameterofappendixmeasuredanteriortoposteriorwithcompre_A,
         Muralthick = Maximalmuralthicknessdistancefromhyperechoicluminalinterfaceto_A,
         fecalith = Presenceoffecalith_A,
         freefluid = Presenceofsurroundingfreefluid_A) %>% 
  mutate(surg_bi = case_when(Surg >=3 ~ 1,
                             Surg <3 ~ 0),
         US2_ = case_when(US2 >=3 ~ 1,
                         US2 <3 ~ 0),
         lossSM = case_when(US2 >= 3 ~ "1",
                            US2 <3 ~ "0"),
         Muralthick = as.numeric(Muralthick)
         ) 
US = data %>% dplyr::select(US2_, surg_bi)
US = drop_na(US) %>% filter()

data$fecalith[data$fecalith == "Missing"] = NA
data$freefluid[data$freefluid == "Missing"] = NA

data = data %>% dplyr::select(surg_bi, Appdiam, Muralthick, fecalith , freefluid , lossSM,Surg)
data = drop_na(data) # 61
str(data) 
```

## Visualization

```{r}
# data import
data = read_delim("data.dat", "\t", escape_double = FALSE, trim_ws = TRUE) 

data = data[1:96,] %>% dplyr::select(SurgPathFindings,SeqentialStaging, Largestdiameterofappendixmeasuredanteriortoposteriorwithcompre_A, Maximalmuralthicknessdistancefromhyperechoicluminalinterfaceto_A,  Presenceoffecalith_A, Presenceofsurroundingfreefluid_A) %>% 
  rename(Surg = SurgPathFindings,
         US2 = SeqentialStaging,
         Appdiam = Largestdiameterofappendixmeasuredanteriortoposteriorwithcompre_A,
         Muralthick = Maximalmuralthicknessdistancefromhyperechoicluminalinterfaceto_A,
         fecalith = Presenceoffecalith_A,
         freefluid = Presenceofsurroundingfreefluid_A) %>% 
  mutate(Muralthick = as.numeric(Muralthick),
         Surg = as.factor(Surg)) %>% 
  filter(!is.na(Surg)) 

data = data[-59,]

data$fecalith[data$fecalith == "Missing"] = NA
data$freefluid[data$freefluid == "Missing"] = NA


# Appdiam
dataA = data %>% filter(Appdiam <= 2)
ggplot(dataA, aes(x = Surg, y = Appdiam)) + geom_boxplot()

aov1 <- aov(Appdiam~Surg, data = data)
summary(aov1)

pairwise.t.test(data$Appdiam, data$Surg, p.adj = 'bonferroni')
TukeyHSD(aov1)

# Muralthick
ggplot(dataA, aes(Surg, y = Muralthick)) + geom_boxplot()

aov2 <- aov(Muralthick~Surg, data = data)
summary(aov2)

pairwise.t.test(data$Muralthick, data$Surg, p.adj = 'bonferroni')
TukeyHSD(aov2)

# Fecalith

ggpiestats(data, x = fecalith, y = Surg)

fisher.test(data$fecalith,data$Surg)

# Free fluid

ggpiestats(data, x = freefluid, y = Surg)

fisher.test(data$freefluid,data$Surg)
```




## Logistic regression


1) loss of normal echogenic submucosal layer: In the dataset, loss of SM layer is captured by the ultrasound and is described as stage IV: Gangrenous.
2) mural hyperemia: Maximum mural thickness 
3) periappendiceal fluid: Presence of surrounding free fluid
4) appendicolith: presence of fecalith
5) appendix wall diameter: largest diameter of appendix

```{r}
# choose model
glm1 = glm(data = data, surg_bi ~ . , family=binomial(link='logit')) 
summary(glm1)

step = glm2 %>% stepAIC(trace = FALSE)
glm2 = glm(data = data, surg_bi ~ Appdiam + fecalith + freefluid , family=binomial(link='logit')) 
summary(glm2)

data_ = data %>% mutate(
  freefluidcom = case_when(freefluid == "Yes; complex" ~ "1",
                           freefluid != "Yes; complex" ~ "0"),
  freefluid_ = case_when(freefluid == "Yes; complex" ~ "1",
                         freefluid == "Yes; simple" ~ "1",
                         freefluid == "Yes;" ~ "1",
                         freefluid == "No" ~ "0")
)

glm3 = glm(data = data_,surg_bi ~ Appdiam + Muralthick +  fecalith + freefluid_ + lossSM , family=binomial(link='logit')) 
summary(glm3)
step3 = glm3 %>% stepAIC(trace = FALSE)
glm3_ = glm(formula = surg_bi ~ Appdiam + lossSM, family = binomial(link = "logit"), 
    data = data_)
summary(glm3_)



glm4 = glm(data = data_, surg_bi ~ Appdiam + Muralthick +  fecalith + freefluidcom + lossSM, family=binomial(link='logit')) 
summary(glm4)
step4 = glm4 %>% stepAIC(trace = FALSE)
glm4_ = glm(formula = surg_bi ~ Appdiam  + freefluidcom + lossSM, family = binomial(link = "logit"), 
    data = data_)
summary(glm4_)


# prediction


confusionMatrix(data = as.factor(US$US2_),
                   reference = as.factor(US$surg_bi), positive = "1")

glm = glm4_
glm_prob = predict(glm, newdata = data_, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                   reference = as.factor(data_$surg_bi), positive = "1")


glm = glm3_
glm_prob = predict(glm, newdata = data_, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
CM = confusionMatrix(data = as.factor(glm_pred),
                   reference = as.factor(data_$surg_bi), positive = "1")


# use appendiam as continuous variable in glm
roc = roc(as.factor(data_$surg_bi), glm_prob)
plot(roc, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc), add = TRUE)

```

### set cut off of appendiam

```{r}
Data = data_
hist(Data$Appdiam)
grid = seq(0.8,1.4, by = 0.1)
diam = list()
for (i in 1:7) {
  diam[[i]] = ifelse(Data$Appdiam > grid[i], 1, 0) 
  diam[[i]] = as.factor(diam[[i]])
  Data = cbind(Data, diam[[i]])
  
}

colnames(Data) = c("surg_bi", "Appdiam", "Muralthick", "fecalith", "freefluid", "lossSM", "freefluidcom", "freefluid_", "diam1", "diam2", "diam3", "diam4", "diam5", "diam6", "diam7")

# diam1
glm3_1 = glm(formula = surg_bi ~ diam1 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_1)

glm = glm3_1
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc1 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc1, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc1), add = TRUE)

# diam2
glm3_2 = glm(formula = surg_bi ~ diam2 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_2)

glm = glm3_2
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc2 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc2, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc2), add = TRUE)

# diam3
glm3_3 = glm(formula = surg_bi ~ diam3 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_3)

glm = glm3_3
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
CM3 = confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc3 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc3, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc3), add = TRUE)

# diam4
glm3_4 = glm(formula = surg_bi ~ diam4 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_4)

glm = glm3_4
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
CM4= confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc4 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc4, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc4), add = TRUE)

# diam5
glm3_5 = glm(formula = surg_bi ~ diam5 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_5)

glm = glm3_5
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
CM5= confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc5 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc5, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc5), add = TRUE)

# diam6
glm3_6 = glm(formula = surg_bi ~ diam6 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_6)

glm = glm3_6
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc6 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc6, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc6), add = TRUE)

# diam7
glm3_7 = glm(formula = surg_bi ~ diam7 + lossSM, family = binomial(link = "logit"), 
    data = Data)
summary(glm3_7)

glm = glm3_7
glm_prob = predict(glm, newdata = Data, type = "response")
glm_pred = rep("0", 61)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(Data$surg_bi), positive = "1")
roc7 = roc(as.factor(Data$surg_bi), glm_prob)
plot(roc7, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc7), add = TRUE)

# pick 3,4,5
CM$overall[1:2]
CM3$overall[1:2]
CM4$overall[1:2]
CM5$overall[1:2]
CM$byClass[1:2]
CM3$byClass[1:2]
CM4$byClass[1:2]
CM5$byClass[1:2]
CM$table
CM3$table
CM4$table
CM5$table


plot(roc, legacy.axes = TRUE,  col = 1, thresholder = "best", print.thres ="best")
plot(roc3, legacy.axes = TRUE, col = 2, add = TRUE)
plot(roc4, legacy.axes = TRUE,  col = 3, add = TRUE)
plot(roc5, legacy.axes = TRUE,  col = 4, add = TRUE)
legend("bottomright",
       legend = c("continuous, auc: 0.882","Cutoff at 1, auc = 0.876","Cutoff at 1.1, auc = 0.883","Cutoff at 1.2, auc = 0.890"),
       col = 1:4, lty = 1
       )

plot(roc, legacy.axes = TRUE,  col = 1, thresholder = "best", print.thres ="best")
plot(roc3, legacy.axes = TRUE, col = 2, thresholder = "best", print.thres ="best")
plot(roc4, legacy.axes = TRUE,  col = 3, thresholder = "best", print.thres ="best")
plot(roc5, legacy.axes = TRUE,  col = 4, thresholder = "best", print.thres ="best")
```


## Ordinal 
```{r}

# data import
data = read_delim("data.dat", "\t", escape_double = FALSE, trim_ws = TRUE) 

# 96 patients and remove comments column
data = data[1:96,] %>% dplyr::select(SurgPathFindings,SeqentialStaging, Largestdiameterofappendixmeasuredanteriortoposteriorwithcompre_A, Maximalmuralthicknessdistancefromhyperechoicluminalinterfaceto_A,  Presenceoffecalith_A, Presenceofsurroundingfreefluid_A) %>% 
  rename(Surg = SurgPathFindings,
         US2 = SeqentialStaging,
         Appdiam = Largestdiameterofappendixmeasuredanteriortoposteriorwithcompre_A,
         Muralthick = Maximalmuralthicknessdistancefromhyperechoicluminalinterfaceto_A,
         fecalith = Presenceoffecalith_A,
         freefluid = Presenceofsurroundingfreefluid_A) %>% 
  mutate(surg_bi = case_when(Surg >=3 ~ 1,
                             Surg <3 ~ 0),
         US2_ = case_when(US2 >=3 ~ 1,
                         US2 <3 ~ 0),
         lossSM = case_when(US2 >= 3 ~ "1",
                            US2 <3 ~ "0"),
         Muralthick = as.numeric(Muralthick),
         freefluid = case_when(freefluid == "Yes" ~"Yes;",
                               freefluid != "Yes" ~ freefluid)
         ) 

data_ = data %>% mutate(
  freefluidcom = case_when(freefluid == "Yes; complex" ~ "1",
                           freefluid != "Yes; complex" ~ "0"),
  freefluid_ = case_when(freefluid == "Yes; complex" ~ "1",
                         freefluid == "Yes; simple" ~ "1",
                         freefluid == "Yes;" ~ "1",
                         freefluid == "No" ~ "0"),
  surg = factor(Surg)
) 

data_ = drop_na(data_)

ord = polr(surg ~ Appdiam + Muralthick + fecalith  +freefluidcom + lossSM, data = data_  )
summary(ord)
```


## Decision tree - pass
```{r}
rpart1 = rpart(surg_bi ~ Appdiam + Muralthick +  fecalith + freefluid_ + lossSM,
   method="class", data=data_)
rpart2 = rpart(surg_bi ~ Appdiam + Muralthick +  fecalith + freefluidcom + lossSM,
   method="class", data=data_)

print(rpart1)
print(rpart2)
rpart.plot(rpart1)
rpart.plot(rpart2)

```

## Supplement 1.13

```{r}
data = data %>% dplyr::select(surg_bi, Appdiam, Muralthick, fecalith , freefluid , lossSM,Surg)

data_ = data %>% mutate(
  freefluid_ = case_when(freefluid == "Yes; complex" ~ "1",
                         freefluid == "Yes; simple" ~ "1",
                         freefluid == "Yes;" ~ "1",
                         freefluid == "No" ~ "0"),
  app_ = case_when(Appdiam >1 ~ "1",
                   Appdiam <=1 ~ "0"),
  fecalith = case_when(fecalith == "Yes" ~ "1",
                       fecalith == "No" ~ "0")
) %>% dplyr::select(surg_bi, app_, fecalith, freefluid_, lossSM) %>% drop_na()#74

glm1 = glm(data = data_, surg_bi ~ . , family=binomial(link='logit')) 
summary(glm1)

glm = glm1
glm_prob = predict(glm, newdata = data_, type = "response")
glm_pred = rep("0", 74)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(data_$surg_bi), positive = "1")
roc1 = roc(as.factor(data_$surg_bi), glm_prob)
plot(roc1, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc1), add = TRUE)

glm2 = glm(data = data_, surg_bi ~ app_ + fecalith + lossSM , family=binomial(link='logit')) 
summary(glm2)

glm = glm2
glm_prob = predict(glm, newdata = data_, type = "response")
glm_pred = rep("0", 74)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(data_$surg_bi), positive = "1")
roc2 = roc(as.factor(data_$surg_bi), glm_prob)
plot(roc2, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc2), add = TRUE)

glm3 = glm(data = data_, surg_bi ~ app_ + freefluid_ + lossSM , family=binomial(link='logit')) 
summary(glm3)

glm = glm3
glm_prob = predict(glm, newdata = data_, type = "response")
glm_pred = rep("0", 74)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(data_$surg_bi), positive = "1")
roc3 = roc(as.factor(data_$surg_bi), glm_prob)
plot(roc3, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc3), add = TRUE)

glm4 = glm(data = data_, surg_bi ~ app_ +  lossSM , family=binomial(link='logit')) 
summary(glm4)

glm = glm4
glm_prob = predict(glm, newdata = data_, type = "response")
glm_pred = rep("0", 74)
glm_pred[glm_prob >0.5] = "1"
confusionMatrix(data = as.factor(glm_pred),
                reference = as.factor(data_$surg_bi), positive = "1")
roc4 = roc(as.factor(data_$surg_bi), glm_prob)
plot(roc4, legacy.axes = TRUE, print.auc = TRUE, col = 4)
plot(smooth(roc4), add = TRUE)


confusionMatrix(data = as.factor(data_$lossSM),
                reference = as.factor(data_$surg_bi), positive = "1")


fisher.test(data_$lossSM, data_$fecalith)

confusionMatrix(data = as.factor(data_$fecalith),
                reference = as.factor(data_$lossSM), positive = "1")


fisher.test(data_$lossSM, data_$freefluid_)

confusionMatrix(data = as.factor(data_$freefluid_),
                reference = as.factor(data_$lossSM), positive = "1")

fisher.test(data_$lossSM, data_$app_)

confusionMatrix(data = as.factor(data_$app_),
                reference = as.factor(data_$lossSM), positive = "1")
```


